---
title: "KadenFlorsheim_A6"
author: "Kaden Florsheim"
date: "11/1/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, warning = F, message = F)
```

```{r}
library(tidyverse)
library(sf)
library(tigris)
library(mapview)
library(leaflet)
library(survey)
library(censusapi)
Sys.setenv(CENSUS_KEY="9fbd5ddd430b595b8f3715733cae2b75c18be92e")
```


```{r}
# pums_2019_1yr <- getCensus(
#   name = "acs/acs1/pums",
#   vintage = 2019,
#   region = "public use microdata area:*",
#   regionin = "state:06",
#   vars = c(
#     "SERIALNO",
#     "SPORDER",
#     "PWGTP",
#     "WGTP",
#     "YBL",
#     "BLD",
#     "TEN",
#     "MV",
#     "HINCP",
#     "AGEP"
#   )
# )
#saveRDS(pums_2019_1yr, "a6_pums.rds")
```

```{r}
pums_2019_1yr <- readRDS("a6_pums.rds")

ca_pumas <-
  pumas("CA", cb = T, progress_bar = F)

sf_boundary <-
  counties("CA", cb = T, progress_bar = F) %>%
  filter(NAME == "San Francisco")

sf_pumas <-
  ca_pumas %>% 
  st_centroid() %>% 
  .[sf_boundary, ] %>% 
  st_drop_geometry() %>% 
  left_join(ca_pumas %>% select(GEOID10)) %>% 
  st_as_sf()
```

```{r}
sf_pums <- pums_2019_1yr %>% 
  mutate(
    PUMA = str_pad(public_use_microdata_area,5,"left","0")
  ) %>% 
  filter(PUMA %in% sf_pumas$PUMACE10)

sf_pums_clean <- sf_pums %>%
  mutate(
    YBL = as.numeric(YBL),
  ) %>% 
  filter(YBL %in% 1:3) %>%
  arrange(as.numeric(AGEP)) %>%
  group_by(SERIALNO) %>% 
  summarize_all(first) %>% 
  mutate(
    AGEP = as.numeric(AGEP),
    HINCP = as.numeric(HINCP),
    leadrisk = ifelse(
      as.numeric(HINCP) <= 90000 &
        as.numeric(AGEP) <= 6,
      1,
      0
    )) %>% 
  mutate(across(
    c(BLD, TEN, YBL, WGTP, MV, HINCP, PUMA),
    factor))
```

```{r}
logit_model <- glm(
  leadrisk ~  BLD + TEN  + MV + PUMA,
  family = quasibinomial(),
  data = sf_pums_clean
)

summary(logit_model)
```

```{r}
sample_predict <- sample_n(sf_pums_clean, 1)

predict(logit_model, sample_predict, type = "response")
```


```{r}
predicted_score <- 
  predict(logit_model, sf_pums_clean, type = "response") 

cbind(sf_pums_clean, predicted_score)
```

```{r}
summary_2x2 <-
  sf_pums_clean %>% 
  mutate(
    leadrisk = ifelse(
      leadrisk == 1, 
      "Yes (Lead risk)", 
      "No (No lead risk)"
    ),
    predicted_leadrisk = ifelse(
      predicted_score >= .1,
      as.numeric(WGTP),
      0
    ),
    predicted_noleadrisk = ifelse(
      predicted_score < .1,
      as.numeric(WGTP),
      0
    ) 
  ) %>% 
  group_by(leadrisk) %>% 
  summarize(
    predicted_leadrisk = sum(predicted_leadrisk),
    predicted_noleadrisk = sum(predicted_noleadrisk)
  )
  
 # pull(leadrisk) %>% 
 # table(predicted_score > 0.1)
```

```{r}
summary_2x2
```

##Households that exceeded the threshold score: BLD- The BLD values generally were either 2 or 3. This suggests that these are typically One-family houses detatched or attatched. TEN- The values were generally either 1 or 3. This suggests that the houses/apartments are either owned with mortgage or loan or they are rented. MV- 2, 3, and 4 were the most common values but there were outliers. This suggests that residents moved into the houses/apartments .13 to 23 months, 2 to 4 years, and 5 to 9 years. The organization would mail out roughly 11 thousand testing kits. Since the many of the homes were single family attatched/detatched, I would expect 80% of the kits to go to low incomes houses with children. Roughly 22 percent of at risk children will recieve kits.One way we can improve our strategy is by sending a postcard to every home the has a BLD value of 2 or sending a postcard to every home with a TEN value of 1 or 3. Using the data that we have we can increase the number of people that get kits.  

```{r}
sf_boundary <-
  counties("CA", cb = T, progress_bar = F) %>%
  filter(NAME == "San Francisco")

ca_pumas <-
  pumas("CA", cb = T, progress_bar = F)

sf_pumas <-
  ca_pumas %>% 
  st_centroid() %>% 
  .[sf_boundary, ] %>% 
  st_set_geometry(NULL) %>% 
  left_join(ca_pumas %>% select(GEOID10)) %>% 
  st_as_sf()
```

```{r}
temp <- tempfile()
download.file("https://www2.census.gov/programs-surveys/acs/data/pums/2019/1-Year/csv_hca.zip",destfile = temp)

pums_wgtp_2019_1yr <- read_csv(unzip(temp,"psam_h06.csv"))

unlink(temp)
```

```{r}
library(readxl)
library(spdep)
library(spatialreg)
```

```{r}
data <- sf_pums_clean %>%
  left_join(pums_wgtp_2019_1yr, by = c("SERIALNO")) %>% 
  select(
    starts_with("WGTP")
  ) %>% 
  select(-WGTP.x,-WGTP.y)
```

```{r}
logit_survey_design <- svrepdesign(
  data = sf_pums_clean,
  type = "ACS",
  repweights = data,
  weights = ~as.numeric(WGTP)
)
```

```{r}
logit_survey_model <- svyglm(
  formula = leadrisk~ + BLD + TEN  + MV + PUMA,
  family = quasibinomial(),
  design = logit_survey_design
)
```

```{r}
summary(logit_survey_model)
```

```{r}
sample_predict <- sample_n(sf_pums_clean, 1)

predict(logit_survey_model, sample_predict, type = "response")
```

```{r}
predicted_score <- 
  predict(logit_survey_model, sf_pums_clean, type = "response") 

cbind(sf_pums_clean, predicted_score)
```

```{r}
summary_2x2 <-
  sf_pums_clean %>% 
  mutate(
    leadrisk = ifelse(
      leadrisk == 1, 
      "Yes (Lead risk)", 
      "No (No lead risk)"
    ),
    predicted_leadrisk = ifelse(
      predicted_score >= .1,
      as.numeric(WGTP),
      0
    ),
    predicted_noleadrisk = ifelse(
      predicted_score < .1,
      as.numeric(WGTP),
      0
    ) 
  ) %>% 
  group_by(leadrisk) %>% 
  summarize(
    predicted_leadrisk = sum(predicted_leadrisk),
    predicted_noleadrisk = sum(predicted_noleadrisk)
  )
```

```{r}
summary_2x2
```
##The number of kits mailed out increased to 13,000 kits and the number of at risk kids that did not receive kits decreased by 286. I would predict a similar success rate for how many kits would end up at target housholds (70%-80%.) The percent of at risk children likly to be reached would increase by 3% to 25%. 